The Error Estimate of Force Calculation in the Heterogeneous Molecular Systems

### 2011-09-27 shortRange.inhomo/writeDown/main.tex
back up, finished intro

\documentclass[aps,pre,preprint]{revtex4}
% \documentclass[aps,jcp,groupedaddress,twocolumn,unsortedaddress]{revtex4}

@@

\title{The Error Estimate of Force Calculation in Heterogeneous and Correlated Molecular Systems}
\author{Han Wang}
\affiliation{LMAM and School of Mathematical
  Sciences, Peking University}
\author{Pingwen Zhang}
% \email{pzhang@pku.edu.cn}
\affiliation{LMAM and School of Mathematical
  Sciences, Peking University}

@@

@@

If a particle with cut-off radius $r_c$ is located at position $\v r$,
we denote the computed force (or cut-offed force) by $\tilde{\v F}(\v
r)$. The difference between the exact force and the computed force,
namely the error force, is denoted by $\Delta\v F(\v r) = \v F(\v r) -
\tilde{\v F}(\v r)$. The error force can be expressed by
\begin{align}
  \Delta\v F(\v r) = \sum_{j}\, \v f^c(\v r - \v r_j)
\end{align}
where $\v f^c$ is the complementary pairwise short-range force defined
by
\begin{align}
  \v f^c(\v r) =
  \left\{
  \begin{array}{ll}
    0, & \vert\v r\vert\leq r_c; \\
    \v f(\v r), & \vert\v r\vert > r_c.
  \end{array}
  \right.
\end{align}
In a homogeneous system with periodic boundary condition applied, the
surounding of any particle is isotropic, so the mean of the error
force vanishes. However, in a heterogenous system, the mean of the
error force is not necessarily zero, but
\begin{align} \label{eqn:tmp3} %\nonumber
  \langle\Delta\v F(\v r)\rangle
  &=
  \big\langle
  \sum_{j}\, \v f^c(\v r - \v r_j)
  \big\rangle 
  =
  \int_{\mathbb R^3}\v f^c(\v r - \v r')\rho(\v r')\,\d d\v r'
\end{align}
Where $\rho(\v r)$ is the particle number density at the position $\v
r$. And notice the error force implicitly depends on the cut-off
radius used for the particle.

The widely accepted definition of the force error is the root mean
square (RMS) error, namely the square root of the second moment of the
error force is $\mathcal E(\v r) = \sqrt{\langle\vert\Delta\v F(\v
  r)\vert^2\rangle}$, which can be calculate by
\begin{align} \nonumber
  \langle\vert\Delta\v F(\v r)\vert^2\rangle
  =\,&
  \big\langle\sum_{j,k}\v f^c(\v r - \v r_j)\cdot\v f^c(\v r - \v r_k)\big\rangle \\ \nonumber
  =\,&
  \big\langle\sum_j\vert\v f^c(\v r - \v r_j)\vert^2\big\rangle +
  \big\langle\sum_{j\neq k}\v f^c(\v r - \v r_j)\cdot\v f^c(\v r - \v r_k)\big\rangle \\ \nonumber
  =\,&
  \int_{\mathbb R^3}\vert\v F^c(\v r - \v r')\vert^2\rho(\v r')\,\d d\v r'
  + \\\label{eqn:tmp4}
  \,&
  \int_{\mathbb R^3\times\mathbb R^3}\v F^c(\v r - \v r')\cdot\v F^c(\v r - \v r'')\rho(\v r', \v r'')\,\d d\v r'\d d\v r'',
\end{align}
The density $\rho(\v r')$ and the pair density $\rho(\v r', \v r'')$
are related to the probability density $p(\v r')$ and pair probability
density $p(\v r', \v r'')$ by
\begin{align}
  \rho(\v r') &= N\,p(\v r') \\
  \rho(\v r', \v r'') &= N(N-1)\,p(\v r', \v r'')
\end{align}
The densities are periodically extented to the whole space $\mathbb
R^3$.
% The problem now this how to discretize the integrals in
% \eqref{eqn:tmp4}. With the piecewise constant numerical integration
% formulus, we have:
Now, let
\begin{align} \nonumber
  \rho(\v r', \v r'')
  & = N (N-1) \,p(\v r', \v r'') \\\nonumber
  & = N (N-1) \big\{ p(\v r')p(\v r'') + [\,p(\v r', \v r'') -  p(\v r')p(\v r'')\,]\big\} \\\nonumber
  &= \frac{N-1}N \rho(\v r')\rho(\v r'') + C (\v r', \v r'')\\ \label{eqn:tmp7}
  &\approx \rho(\v r')\rho(\v r'') + C (\v r', \v r'')
\end{align}
the function
\begin{align}
C (\v r', \v r'') = N (N-1) [\,p(\v r', \v r'') -  p(\v r')p(\v r'')\,]
\end{align}
denotes the correlation between
position $\v r'$ and $\v r''$. Therefore, \eqref{eqn:tmp4} becomes
\begin{align} \nonumber
  \langle\vert\Delta\v F(\v r)\vert^2\rangle
  = &\,
  \int_{\mathbb R^3}\vert\v f^c(\v r - \v r')\vert^2\rho(\v r')\,\d d\v r' \,+ \\\nonumber
  &\,
  \bigg[\int_{\mathbb R^3}\v f^c(\v r - \v r')\rho(\v r')\,\d d\v r'\,\bigg]^2 + \\\nonumber
  &\,
  \int_{\mathbb R^3\times\mathbb R^3}\v f^c(\v r - \v r')\cdot\v f^c(\v r - \v r'')\,C(\v r', \v r'')\,\d d\v r'\d d\v r'' \\\label{eqn:tmp5}
  = &\,
  \mathcal E_{\textrm{homo}}(\v r) +
%  \langle\Delta\v F(\v r)\,\rangle^2 +
  \mathcal E_{\textrm{hetero}}(\v r) +
  \mathcal E_{\textrm{correlation}}(\v r)
\end{align}
If the system is homogeneous and the correlation between particles is
neglected, then only the first term, which is called the homogeneity
error, is left in \eqref{eqn:tmp5}. This term is originated from the
flucuation of the error force $\Delta\v F(\v r)$. The second term is
called the heterogeneity error, because it only exists when the system
is heterogenous. It is worth poining out that the heterogeneity error
is the squared mean of the error force, namely $\mathcal
E_{\textrm{hetero}}(\v r) = \langle\Delta\v F(\v r)\,\rangle^2$.  The
last term in \eqref{eqn:tmp5} originates from the correlation between
the particles, so it is called the correlation error.  The correlation
error involves a double integrals, which implies that this term
requires more compuational cost than the homogeneity and heterogeneity
errors. The rest of this paper assumes the system is far from the
critical point, therefore the correlation error does not dominate in the
system, and can be savely neglected.

\subsection{The ways to reduce force error in a heterogenous system}

In Fig. \ref{fig:tmp1}, the testing system containing $N=16000$
standard Lennard-Jone particles is set up in a tube with periodic
boundary condition. The liquid droplet (in the middle of the tube) is
in equilibrium with the surrounding vapor environment. The red line in
the Fig. \ref{fig:tmp2} denotes the density profile (on direction $x$)
of this heterogeneous system. The green square denotes the real RMS
force error distribution. It is clear that the error form two peaks in
the interfacial region, which is much higher than the relative
constant error of the bulk liquid and the bulk vapor. The dashed and
the dotted blue line are the estimated homogeneity and heterogeneity
errors, and the solid blue line is the sum. In the bulk region, the
homogeneity error dominates while in the interfacial region, the
heterogeneity error dominates. The esitmated error matches perfactly
with the real error at the interfacial region, but presents
overestimation in the bulk regions. The overestimation probably stems
from the neglection of the correlation error in \eqref{eqn:tmp5}.

Traditionally, a very large cutoff radius is used to reach precise
simulation results. Since heterogeneity error dominate the force error
of the system, such a big cutoff radius is a waste in the bulk liquid
and vapor regions. One possible way is to use adapt cut-off radius
according to the error distribution, so that larger cut-off radii are
used in the interfacial region while smaller cut-off radii are used in
the bulk regions.  More precisely, the aim of the adaptive cut-off
method is to equivalently distributed the force error over the
simulation region.

The other way is to screen out the heterogeneity error by a applying a
correction force.  Starting from equation \eqref{eqn:tmp5}, the
variation of the error force can be calculated by
\begin{align} \nonumber
  \textrm{Var}[\Delta\v F(\v r)]
  & =
  \big\langle\,
  [\,\Delta\v F(\v r) - \langle\Delta\v F(\v r)\rangle\,]^2
  \big\rangle \\\nonumber
  & =
  \langle\vert\Delta\v F(\v r)\vert^2\rangle -
  \langle\Delta\v F(\v r)\,\rangle^2 \\
  & =
  \mathcal E_{\textrm{homo}}(\v r) +
  \mathcal E_{\textrm{corr}}(\v r)  
\end{align}
This means if we correct the computed force by the mean of the error
force, namely applying to each particle $\tilde{\v F}^c(\v r) =
\tilde{\v F}(\v r) - \langle\Delta\v F(\v r)\rangle$ instead of
$\tilde{\v F}(\v r)$. The error of this corrected force is reduced to
$\mathcal E_{\textrm{homo}}(\v r) + \mathcal E_{\textrm{corr}}(\v r)$
and contains no heterogeneous contribution.  These two possibilities
will be addressed and tested later in this paper.

@@

The liquid-vapor coexistence simulation is a very typical
heterogeneous molecular system, which is widely used to study the
phase coexistence properties, for example, the equilibrium liquid and
vapor density, and the surface tension. The system contains 16,000
standard Lennard-Jones particles that saperated into liquid phase in
the center of the box and vapor phase surrounded. The size of the
simulation box is $L_x^\ast \times \L_y^\ast \times L_z^\ast =
150\times 21\times 21$. This simulation box permit a maximum cut-off
radius of $r_c^\ast = 10$, and eliminate the finite scale effect
discussed in literatures \recheck {give citations here}. In all
simulations, the maximum cut-off radius considered is $10$, but the
reference (precise value) results are given by a simulation carried
out in a 24,000 particles system with simulation box $L_x^\ast \times
\L_y^\ast \times L_z^\ast = 150\times 27\times 27$, and a uniform
cut-off radius of $r_c^\ast = 13$.

Figure \ref{fig:tmp2} presents the real RMS force error (in green
square) in a system using a uniform cut-off raidus of $r_c^\ast =
7.5$. The number density distribution along $x$ direction that is
shown by the solid red line for reference.  The error is comparatively
constant at the bulk liquid and gas region and two paaks, which are
more than one order of maganitude higher than the bulk force errors,
form at the interfacial region.  The estimated RMS force error is
presented in solid blue line.  The estimate follows the real error
very well at the interfacial region, but slightly over estimate the
error in the bulk regions.  Two component of interest, namely the
homogeneous error and heterogeneous error are presented in
Fig. \ref{fig:tmp2} by dashed and dotted blue lines, respectively.  In
the bulk regions, the dashed blue line overlaps with the solid blue
line, that means the homogeneous error dominates.  In the interfacial
region, in the contrary, the heterogeneous error dominate, because the
dotted blue line overlaps with the solid one.

>>>>

\documentclass[aps,pre,preprint]{revtex4}
% \documentclass[aps,pre,twocolumn]{revtex4}
% \documentclass[aps,jcp,groupedaddress,twocolumn,unsortedaddress]{revtex4}

@@

\title{The Error Estimate of Force Calculation in the Heterogeneous Molecular Systems}
\author{Han Wang}
\affiliation{LMAM and School of Mathematical
  Sciences, Peking University}
\author{Pingwen Zhang}
% \email{pzhang@pku.edu.cn}
\affiliation{LMAM and School of Mathematical
  Sciences, Peking University}

@@

\section{Introduction}
Non-bonded short-range interaction is encountered in nearly every
molecular simulation.  A naive idea to calculate short-range
interactions is to explicitly calculate all pairwise
interactions. This results in a computational cost scaling $\mathcal
O(N^2)$ per time step, which rapidly becomes inefficient as the number
of particles $N$ grows large. A better way is to introduce a cut-off
radius, outside of which all pairwise interactions are simply
neglected. In combination with the cell list and the neighbor list
algorithms \cite{frenkel02b}, the total computational cost of the
short-range interactions can be reduced to an acceptable level of
$\mathcal O(N)$. The cut-off radius is only a user provided
algorithmetic parameter, so it should have no effect on the
physical/chemical properties of the system.  Therefore, it is very
important to test the convergence of the simulation results with
respect to the used cut-off radius before collecting data and deriving
conclusions. Fortunately, in most homogeneous systems, a satisfactory
convergence can be achieved even with a very small cut-off value, for
example, $2.5$ -- $4.0\sigma$.  Moreover, it is possible to correct
the systematic error of the potential energy, the pressure and the
free energy by applying the standard long range correction
(LRC)~\cite{allen87a}, which assumes the uniformity outside the
cut-off and integrates the contribution of the mentioned properties to
infinity.  However,
% The spacial inhomogenety of the system in terms of particle density,
% particle type and change distribution is very common in natrual
% system.
the direct molecular simulation of a inhomogeneous system will
encounter non-trival difficulty that does not exist in the homogeneous
bulk simulations.  The measured properties are found to depend on the
cut-off radius even at a very large value~\cite{duque2004some}. Also
the contiuity at the cut-off and the long-range correction involved
may have foundamental effect on the simulation
results~\cite{trokhymchuk1999computer, shen2007comparative,
  guo1997long, mecke1997molecular, janecek2006long, goujon2004monte}.
Therefore, it is necessary to systemetically study the precision of
short-range interaction calculation, and to develop methods screening
out the aforementioned cut-off artifect in the inhomogeneous systems.

% In this paper, we focus on the systems in which
% atoms/particles are interacting via short-range interactions, because
% it is a good starting point of the study of more complex systems, for
% example, the systems with long-range electrostatic interactions.  The
% simulation results in the inhomogeneous systems are found to depend on
% a number of factors of treating the short-range interaction in the
% simulation. For example, the way of calculating the short-range
% interaction (simple cut-off method or Ewald
% summation)~\cite{karasawa1989acceleration, lopez2002thermodynamic,
%   lopez2006liquid, shi2006molecular, ismail2007application}, the
% cut-off radius~\cite{trokhymchuk1999computer, duque2004some,
%   shen2007comparative}, the way of long-range
% correction~\cite{guo1997long, mecke1997molecular, janecek2006long,
%   goujon2004monte} and the size and shape of the simulation
% box~\cite{chen1995area, orea2005oscillatory, biscay2009calculation}.
% It should be noticed that the physical properties are the nature of
% the system, and should NOT depend on the details of the interaction
% calculation.

A promising way to quantitatively analyze the undesirable cut-off
effects is to express these artifacts in terms of the difference
between the cut-offed interaction and the exact interaction, namely
the error. In homogeneous systems, the error analysis of the
short-range interaction has been well established,
see~\cite{kolafa1992cutoff} for example. Also the error
estimates~\cite{hummer1995numerical, kolafa1992cutoff,
  petersen1995accuracy, deserno1998mue2, wang2010optimizing} for
long-range Ewald family methods~\cite{ewald1921die,
  hockney1988computer, deserno1998mue1, darden1993pme, essmann1995spm}
have provided a profound view of the precision of force calculation,
and introduced parameter tuning algorithms~\cite{limbach06a,
  wang2010optimizing} that boost the efficiency of the computation
with a good control of the error. However, in inhomogeneous systems
the error study is still scarce, even for the simplest short-range
interaction.

In this work, we develop the error estimate of the short-range force
in inhomogeneous systems. The force error can be split into the
homogeneous and inhomogeneous part regardless the correlation in the
system. The inhomogeneous error that can be more than one order of
maganitude larger than the homogeneous error plays a dominant role in
the interfacial region. An adaptive cut-off method is hereby proposed
to equally distribte the error according to the error estimate, so
that a substantial computational effort can be saved while the
precision of simulation is preserved. We also proposed a long-range
force correction method based on the error estimate to screen out the
contribution of the inhomogeneous error, which can tremendously
improve the precision of simulation. To show the effectiveness of the
error estimate and the proposed methods, two systems are tested: the
liquid-vapor equilibrium for equilibrium properties, and the colliding
nanoscale clusters for dynamical properties. Finally, a conclusion is
given in section \ref{sec:conclusion}.

% The error estimate of long-range algorithms, for example the Ewald
% family method~\cite{ewald1921die, hockney1988computer,
%   deserno1998mue1, darden1993pme, essmann1995spm} and the fast
% multipole method~\cite{greengard1987fast}, has been well developped in
% the past 20 years ~\cite{hummer1995numerical, kolafa1992cutoff,
%   petersen1995accuracy, deserno1998mue2, wang2010optimizing,
%   dachsel2009error}.

@@

Suppose the system is composed by $N$ identical particles located at
$\v r_1, \v r_2, \cdots, \v r_N$ with periodic boundary condition.
Since the error estimate developed for the one-component system can be
directly extended to multi-component system without substantial
difficulty, the discussion in this paper will focus on one-component
system for symplicity.  These particles are interacting via a
short-range pairwise interaction denoted by $u(\v r)$ and correspond
force denoted by $\v f(\v r)$, with a cut-off radius $r_c > 0$.  This
means the interaction between any pair of particles falling out of the
cut-off radius is neglected.  An infinite cut-off radius will lead to
a precise calculation of interaction, but it is impractical.  The
error of computing interaction is mainly introduced by a finit value
of the cut-off radius. For convenience, the complementary of the
cut-offed short-range force is defined by
\begin{align}
  \v f^c(\v r) =
  \left\{
  \begin{array}{ll}
    0, & \vert\v r\vert\leq r_c; \\
    \v f(\v r), & \vert\v r\vert > r_c.
  \end{array}
  \right.
\end{align}
To measure the error of force computing, we add a testing particle at
position $\v r$. The exact force and the computed force (or cut-offed
force) exerted by the system on the testing particle are denoted by
${\v F}(\v r)$ and $\tilde{\v F}(\v r)$, respectively.  The difference
between them, namely the \emph{error force}, is denoted by $\Delta\v
F(\v r) = \v F(\v r) - \tilde{\v F}(\v r)$, which can be expressed by
\begin{align}
  \Delta\v F(\v r) = \sum_{\v n}\sum_{j}\, \v f^c(\v r - \v r_j + \v n).
\end{align}
Where $\v n$ is simulation box vector, so the summation over $\v n$
indicates that all periodic images are consider in the force
calculation. In a homogeneous system with periodic boundary condition
applied, the surounding of any particle is isotropic, so the
\emph{mean error force} vanishes, namely $\langle\Delta\v F(\v
r)\rangle = 0$. However, in a heterogenous system, the mean error
force is not necessarily zero but
\begin{align} \label{eqn:tmp3} %\nonumber
  \langle\Delta\v F(\v r)\rangle
  &=
  \big\langle
  \sum_{\v n}\sum_{j}\, \v f^c(\v r - \v r_j + \v n)
  \big\rangle 
  =
  \int_{\mathbb R^3}\v f^c(\v r - \v r')\rho(\v r')\,\d d\v r'
\end{align}
Where $\rho(\v r)$ denotes the particle number density at the position
$\v r$, and is periodically extended to $\mathbb R^3$. And notice the
error force implicitly depends on the cut-off radius used for the
interaction.

The widely accepted definition of the force error is the \emph{root mean
square} (RMS) \emph{error}, namely the square root of the second moment of the
error force is $\mathcal E(\v r) = \sqrt{\langle\vert\Delta\v F(\v
  r)\vert^2\rangle}$, which can be calculate by
\begin{align} \nonumber
  \langle\vert\Delta\v F(\v r)\vert^2\rangle
  =\,&
  \big\langle\sum_{j,k}\v f^c(\v r - \v r_j)\cdot\v f^c(\v r - \v r_k)\big\rangle \\ \nonumber
  =\,&
  \big\langle\sum_j\vert\v f^c(\v r - \v r_j)\vert^2\big\rangle +
  \big\langle\sum_{j\neq k}\v f^c(\v r - \v r_j)\cdot\v f^c(\v r - \v r_k)\big\rangle \\ \nonumber
  =\,&
  \int_{\mathbb R^3}\vert\v F^c(\v r - \v r')\vert^2\rho(\v r')\,\d d\v r'
  + \\\label{eqn:tmp4}
  \,&
  \int_{\mathbb R^3\times\mathbb R^3}\v F^c(\v r - \v r')\cdot\v F^c(\v r - \v r'')\rho(\v r', \v r'')\,\d d\v r'\d d\v r'',
\end{align}
The density $\rho(\v r')$ and the pair density $\rho(\v r', \v r'')$
are related to the probability density $p(\v r')$ and pair probability
density $p(\v r', \v r'')$ by
\begin{align}
  \rho(\v r') &= N\,p(\v r') \\
  \rho(\v r', \v r'') &= N(N-1)\,p(\v r', \v r'')
\end{align}
The pair density is also periodically extented to the whole space $\mathbb
R^3$.
% The problem now this how to discretize the integrals in
% \eqref{eqn:tmp4}. With the piecewise constant numerical integration
% formulus, we have:
Now, let
\begin{align} \nonumber
  \rho(\v r', \v r'')
  & = N (N-1) \,p(\v r', \v r'') \\\nonumber
  & = N (N-1) \big\{ p(\v r')p(\v r'') + [\,p(\v r', \v r'') -  p(\v r')p(\v r'')\,]\big\} \\\nonumber
  &= \frac{N-1}N \rho(\v r')\rho(\v r'') + C (\v r', \v r'')\\ \label{eqn:tmp7}
  &\approx \rho(\v r')\rho(\v r'') + C (\v r', \v r'')
\end{align}
the function
\begin{align}
C (\v r', \v r'') = N (N-1) [\,p(\v r', \v r'') -  p(\v r')p(\v r'')\,]
\end{align}
denotes the correlation between
position $\v r'$ and $\v r''$. Therefore, \eqref{eqn:tmp4} becomes
\begin{align} \nonumber
  \langle\vert\Delta\v F(\v r)\vert^2\rangle
  = &\,
  \int_{\mathbb R^3}\vert\v f^c(\v r - \v r')\vert^2\rho(\v r')\,\d d\v r' \,+ \\\nonumber
  &\,
  \bigg[\int_{\mathbb R^3}\v f^c(\v r - \v r')\rho(\v r')\,\d d\v r'\,\bigg]^2 + \\\nonumber
  &\,
  \int_{\mathbb R^3\times\mathbb R^3}\v f^c(\v r - \v r')\cdot\v f^c(\v r - \v r'')\,C(\v r', \v r'')\,\d d\v r'\d d\v r'' \\\label{eqn:tmp5}
  = &\,
  \mathcal E_{\textrm{homo}}(\v r) +
%  \langle\Delta\v F(\v r)\,\rangle^2 +
  \mathcal E_{\textrm{hetero}}(\v r) +
  \mathcal E_{\textrm{correlation}}(\v r)
\end{align}
If the system is homogeneous and the correlation between particles is
neglected, then only the first term, which is called the homogeneity
error, is left in \eqref{eqn:tmp5}. This term is originated from the
flucuation of the error force $\Delta\v F(\v r)$. The second term is
called the heterogeneity error, because it generates due to the
density heterogeneity of the system. It is worth poining out that the
heterogeneity error is the nothing but the squared mean error force,
namely $\mathcal E_{\textrm{hetero}}(\v r) = \langle\Delta\v F(\v
r)\,\rangle^2$.  The last term in \eqref{eqn:tmp5} originates from the
correlation between the particles, so it is called the correlation
error.  The correlation error involves a double integral, which
implies that calculating this term requires more compuational cost
than the homogeneity and heterogeneity errors. The rest of this paper
assumes the system is far from the critical point, therefore the
correlation error does not dominate in the system, and can be savely
neglected.

The naive calculation of the error estimate \eqref{eqn:tmp5} requires
the compuational cost of $\mathcal O(N^2)$. To apply the error
estimate on-the-fly with simulation, the compuational cost should be
reduced to at least $\mathcal O(N\log N)$.  This is achieved by
calculating the convolutions in \eqref{eqn:tmp5} with the fast Fourier
transform. We assume the simulation box is uniformly divided in to
small bins.  The positions of these bins are
\begin{align}
  \v r_{i_1, i_2, i_3} & =
  \frac{i_1}{K_1}\v a_1 + 
  \frac{i_2}{K_2}\v a_2 + 
  \frac{i_3}{K_3}\v a_3,\qquad 0\leq i_\alpha< K_\alpha
\end{align}
where $\v a_\alpha$ are box vectors, and $K_\alpha$ is the number of
division on each direction. In the reciprocal space, a corresponding
lettice is set up:
\begin{align}
  \v k_{k_1, k_2, k_3} & =
  k_1\v a_1^\ast +
  k_2\v a_2^\ast +
  k_3\v a_3^\ast, \qquad 0\leq k_\alpha<K_\alpha
\end{align}
where $\v a_\alpha^\ast$ are reciprocal box vectors defined by $\v
a_\alpha\cdot\v a_\beta^\ast = \delta_{\alpha\beta}$.  In each bin the
error and the particles number density are assumed to be constant,
therefore,
\begin{align}\nonumber
  \mathcal E_{\textrm{homo}}(\v r_{i_1,i_2,i_3})
  & \approx
  \frac1V\sum_{k_1=-\infty}^\infty\sum_{k_2=-\infty}^\infty\sum_{k_3=-\infty}^\infty
  \hat{\mathcal E}_{\textrm{homo}}(\v k_{k_1,k_2,k_3})
  % e^{2\pi i\v k_{k_1,k_2,k_3}\cdot\v r_{i_1,i_2,i_3}}
  \exp\bigg[
  2\pi i\bigg(
  \frac{k_1i_1}{K_1} + \frac{k_2i_2}{K_2} + \frac{k_3i_3}{K_3}
  \bigg)
  \bigg] \\\label{eqn:tmp13}
  & \approx
  \frac1V\sum_{k_1=0}^{K_1}\sum_{k_2=0}^{K_2}\sum_{k_3=0}^{K_3}
  \hat{K}_{\textrm{homo}}(\v k'_{k_1,k_2,k_3})\:
  \hat{\rho}(\v k_{k_1,k_2,k_3})
  % e^{2\pi i\v k_{k_1,k_2,k_3}\cdot\v r_{i_1,i_2,i_3}}
  \exp\bigg[
  2\pi i\bigg(
  \frac{k_1i_1}{K_1} + \frac{k_2i_2}{K_2} + \frac{k_3i_3}{K_3}
  \bigg)
  \bigg]   
\end{align}
here we denote the convolution kernel $\vert\v f^c(\v r)\vert^2$ by
$K_{\textrm{homo}}(\v r)$.  Equation \eqref{eqn:tmp13} simply utilizes
the identity: the Fourier transform of the convolution
$K_{\textrm{homo}}\ast\rho$ is equal to the multiply of
$\hat{K}_{\textrm{homo}}$ and $\hat\rho$.  The prime on $\v
k_{k_1,k_2,k_3}$ means that we should use the periodic image $k_\alpha
- K_\alpha$ instead of $k_\alpha$ when $k_\alpha \geq K_\alpha/2$. By
definition, the Fourier transform of the density is
\begin{align}
  \hat\rho(\v k_{k_1,k_2,k_3})
  & \approx
  \frac{V}{K_1K_2K_3}
  \sum_{i_1=0}^{K_1}\sum_{i_2=0}^{K_2}\sum_{i_2=0}^{K_2}
  \rho(\v r_{i_1,i_2,i_3})
  \exp\bigg[
  -2\pi i\bigg(
  \frac{k_1i_1}{K_1} + \frac{k_2i_2}{K_2} + \frac{k_3i_3}{K_3}
  \bigg)
  \bigg]
\end{align}
The Fourier transform of the homogeneity kernel is 
\begin{align}
  \hat K_{\textrm{homo}}(\v k)
  &=
  \int_{\mathbb R^3}K_{\textrm{homo}}(\v r)e^{-2\pi i\v k\cdot\v r}\,\d d\v r
\end{align}
Most of the short-range interactions are isotropic, therefore,
$K_{\textrm{homo}}(\v r) = K_{\textrm{homo}}(r)$, and
\begin{align}
  \hat{K}_{\textrm{homo}}(\v k)
  =
  \int_{\mathbb R^3}K_{\textrm{homo}}(r)e^{-2\pi i\v k\cdot\v r}\d d\v r
  =
  \int_{r_c}^\infty \frac{2 r}k\, K_{\textrm{homo}}(r) \sin(2\pi kr)\,\d dr
\end{align}

Similarly, by \eqref{eqn:tmp3}, the Fourier mode of the the mean error
force is
\begin{align}\label{eqn:tmp17}
  \langle\Delta\v F\rangle^\wedge(\v k) =
  \hat{\v f}^c(\v k)\,
  \hat\rho(\v k)
\end{align}
where 
\begin{align}\label{eqn:tmp18}
  \hat{\v f}^c(\v k) 
  & = 
  2\pi\frac{\v k}k\, i\,
  \bigg\{
  r_c^2\,u(r_c)
  \bigg[\,
  \frac{2\cos(2\pi kr)}{2\pi kr}
  -\frac{2\sin(2\pi kr)}{(2\pi kr)^2}
  \,\bigg] 
  -
  \int_{r_c}^\infty 2r\,u(r)\sin(2\pi kr)\,\d d\v r
  \bigg\}
\end{align}

\subsection{The adaptive cut-off radius method}

% In Fig. \ref{fig:tmp1}, the testing system containing $N=16000$
% standard Lennard-Jone particles is set up in a tube with periodic
% boundary condition. The liquid droplet (in the middle of the tube) is
% in equilibrium with the surrounding vapor environment. The red line in
% the Fig. \ref{fig:tmp2} denotes the density profile (on direction $x$)
% of this heterogeneous system. The green square denotes the real RMS
% force error distribution. It is clear that the error form two peaks in
% the interfacial region, which is much higher than the relative
% constant error of the bulk liquid and the bulk vapor. The dashed and
% the dotted blue line are the estimated homogeneity and heterogeneity
% errors, and the solid blue line is the sum. In the bulk region, the
% homogeneity error dominates while in the interfacial region, the
% heterogeneity error dominates. The esitmated error matches perfactly
% with the real error at the interfacial region, but presents
% overestimation in the bulk regions. The overestimation probably stems
% from the neglection of the correlation error in \eqref{eqn:tmp5}.

A basic observation in a heterogeneous system is that the RMS force
error is not uniformly distributed. The difference between the maximum
and minimum error can be as large as two orders of maganitude, see
Fig. \ref{fig:tmp2}. Traditionally, a very large cut-off radius is
used to reduce the maximum error to an acceptable level and to reach
convergent simulation results. However, a huge amount of compuational
effort is wasted in the low-error region. The idea of the adaptive
cut-off radius simulation is to use larger cut-off radius for
particles in the high-error region to control the force error, while
to use smaller cut-off radius for particles in the low-error region to
save compuational cost.  The vaidity of this method relies on a
diliberated cut-off choice for each particle so that the force error
is uniformly distributed across the simulation region.  Unfortunately,
the compuational expense of a naive implementation of this method is
at least $\mathcal O(N^2\log N)$.  Becasue to determine the cut-off
radius, we should estimate the RMS force error at least once for each
particle, and the computational cost of each force error estimate is
$\mathcal O(N \log N)$.

% Since heterogeneity error dominate the
% force error of the system, such a big cutoff radius is a waste in the
% bulk liquid and vapor regions. One possible way is to use adapt
% cut-off radius according to the error distribution, so that larger
% cut-off radii are used in the interfacial region while smaller cut-off
% radii are used in the bulk regions.  More precisely, the aim of the
% adaptive cut-off method is to equivalently distributed the force error
% over the simulation region.

To make the adaptive cut-off method feasible,
\begin{enumerate}
\item The space is divided in to small bins, the size of which is the
  same as those mentioned in the section \ref{sec:tmp1.3}.  The
  particles falling in the same bin are assigned by the same cut-off
  radius.
  % The cut-off
  % radius is assumed to be constant within a bin, no matter the
  % position of the partile.
\item The cut-off radius can only be chosen from a discrete series of
  increasing values between a maximum cut-off radius
  $r_c^{\textrm{max}}$ and a minimum cut-off radius
  $r_c^{\textrm{min}}$: $\{\,r_c^{\textrm{min}} = r_c^0 < r_c^1 <
  \cdots < r_c^{M-1} <r_c^M = r_c^{\textrm{max}}\}$.  We call it the
  \emph{candidate set} of cut-off radii.  For convenience but not
  necessarily, the interval between two neighboring cut-off radius in
  uniform, which means
  \begin{align}
    r_c^m = r_c^0 + r_c^{\textrm{step}} m, \quad 0 \leq m \leq M
  \end{align}
  For example, a possible candidate set is $\{2.5\,\sigma,\
  3.0\,\sigma,\ 3.5\,\sigma,\ 4.0\,\sigma,\ \cdots,\ 10.0\,\sigma\}$,
  with $r_c^{\textrm{max}} = 10.0\sigma$, $r_c^{\textrm{min}} =
  2.5\sigma$ and $r_c^{\textrm{step}} = 0.5 \sigma$.
\end{enumerate}
For each cut-off radius in the candidate set, the force error
distribution $\mathcal E(\v r)$ is calculated by estimate
\eqref{eqn:tmp5}. Notice the error is calculated on-the-fly by the
fast algorithm described in section \ref{sec:tmp1.3}, therefore it is
defined on the bins.  For a predetermined error control level
$\mathcal E_{\textrm{C}}$, a satisfactory cut-off radius of a bin is
choosen such that it is the smallest one that satisfying the precision
constrain $\mathcal E(\v r_{i_1,i_2,i_3}) \leq \mathcal
E_{\textrm{C}}$. The computational cost of this algorithm is $\mathcal
O(MN\log N)$.  The idea of this algorithm is based on the observation
that the cut-off radius may not be calculated at a very high
precision, a slightly (and at most $r_c^{\textrm{step}}$) larger
cut-off radius will not waste too much computational cost.  Due to the
thermodynamic flucuation of the system density, the resulting cut-off
distribution $r_c(\v r_{i_1,i_2,i_3})$ (as a function of lattice
position) needs to be refined by the following scheme:
\begin{align}\label{eqn:rc-refine}
  r^{(1)}_c(\v r_{i_1,i_2,i_3}) = \max_{j_\alpha\in I}\,r_c(\v r_{i_1+j_1,i_2+j_2,i_3+j_3}),
  \quad I = \{-1, 0, 1\},
\end{align}
to catch the fluctuating interfacial region.

@@

The liquid-vapor coexistence simulation is a very typical
heterogeneous molecular system, which is widely used to study the
phase coexistence properties, for example, the equilibrium liquid and
vapor density, and the surface tension. The system contains 16,000
standard Lennard-Jones particles that saperated into liquid phase in
the center of the box and vapor phase surrounded.  The conventional
dimensionless unit system is employed: the unit of length, energy,
mass and time are denoted by $\epsilon$, $\sigma$, $m$ and $\tau$
($\tau = \sigma\sqrt{m/\varepsilon}$), respectively.  All quantities
are written in the unitless form by adding the superscript ``$\ast$''.
For example, $r^\ast = r / \sigma$, $T^\ast = k_BT / \epsilon$ and
$P^\ast = P \sigma^3 / \epsilon$.  The MD time step {is} $\Delta
t^\ast = 0.005$. The simulations last for $3\times 10^6$ time
steps. The first $1\times 10^6$ steps are discarded. The quantities of
interest are sampled every 100 time steps. The commonly used blocking
average method~\cite{flyvbjerg1989error} is applied to estimate the
statistical uncertainty of the autocorrelated data. The NVT ensemble
is generated by the Nos\`e-Hoover thermostat~\cite{nose1984molecular,
  hoover1985canonical}.  The size of the simulation box is $L_x^\ast
\times L_y^\ast \times L_z^\ast = 150\times 21\times 21$, as shown in
Fig. \ref{fig:tmp1}. This simulation box permit a maximum cut-off
radius of $r_c^\ast = 10$, and eliminate the finite scale effect
discussed in literatures \cite{orea2005oscillatory,
  biscay2009calculation}. In all simulations, the maximum cut-off
radius considered is $10$, but the reference (precise value) results
are given by a simulation carried out in a 24,000 particles system
with simulation box $L_x^\ast \times L_y^\ast \times L_z^\ast =
150\times 27\times 27$, and a uniform cut-off radius of $r_c^\ast =
13$.
